Alex's Notes and Text Fragments for the JSS Manuscript

### 2021-06-15 paper/paper.Rmd
small edits, esp. coord.-based clustering

"Buffering" refers to the idea of removing observations between test and training set by drawing a circular buffer around one or more observations [@rest2014, @brenning2012, @blockCV].
By introducing more distance between observations, the influence of STAC should be reduced.
Buffering mostly happens in the coordinate space due to the scattered appearance of observations in this space.
\pkg{mlr3spatiotempcv} supports two "buffering" implementations: (1) "Spatial buffering" from package \pkg{blockCV} and (2) "Leave-one-disc-out cross-validation" from \pkg{sperrorest}.

@@

<!--- FIXME: Why refer to the blockCV package? This is a general method, as described (or at least outlined) above. If code from blockCV was used, when this should be acknowledged, but otherwise I don't understand why this method is "from" the blockCV package. -->

<!--- P: This addresses specifically the "buffering" function in the blockCV package which falls into the "buffering" category, the same applies to "spatial disc" from {sperrorest}.
Is this a potential misunderstanding with the overarching level 3 header "Buffering"? -->

The "buffering" method from the \pkg{blockCV} package is conceptually similar to a leave-one-out (LOO) cross-validation approach [@stone1974].
To reduce the similarity between training and test sets, a circular buffer around the observations representing the test set is drawn [@rest2014, @blockCV].
Within this buffer zone, all observations are removed from the training set, thereby reducing the similarity between training and test set.

The implementation in the \pkg{blockCV} package comes with optional support for presence/absence data handling, as often used in the field of species distribution modeling (SDM) [@hijmans2013].
However, by default, the method considers all points for training/test set creation and does not apply any specific SDM adjustments to the data.

@@

The "Spatial CV" method after [@brenning2012] and implemented in package \pkg{sperrorest} is similar to method "environmental blocking" in that it also uses *k-means* clustering for fold creation.
However, it does not make use of arbitrary features but uses the coordinates of all observations to create clusters in the spatial domain.
This is what lead to the naming of `spcv_coords` in \pkg{mlr3spatiotempcv}.

```{r coords, fig.cap="Method 'Spatial CV' showing a k-means clustering approach based on coordinates.", out.width="50%", fig.pos="ht"}
resampling_coords = rsmp("spcv_coords", folds = 5)$instantiate(task)

>>>>

"Buffering" refers to the idea of removing observations between test and training set by drawing a circular buffer around one or more observations [@rest2014, @brenning2012, @blockCV].
By introducing more distance between observations, the influence of STAC should be reduced.
Buffering is mostly applied in the coordinate space due to the scattered appearance of observations in this space.
\pkg{mlr3spatiotempcv} supports two "buffering" implementations: (1) "Spatial buffering" from package \pkg{blockCV} and (2) "Leave-one-disc-out cross-validation" from \pkg{sperrorest}.

@@

<!-- A: This paragraph is a general description of the approach, not a description of unique characteristics of this implementation. General approach of buffering is (and should be) described in the introdoctory paragraph of the Buffering section.

The "buffering" method from the \pkg{blockCV} package is conceptually similar to a leave-one-out (LOO) cross-validation approach [@stone1974].
To reduce the similarity between training and test sets, a circular buffer around the observations representing the test set is drawn [@rest2014, @blockCV].
Within this buffer zone, all observations are removed from the training set, thereby reducing the similarity between training and test set.
-->

The implementation in the \pkg{blockCV} package comes with optional support for presence/absence data handling, as often used in the field of species distribution modeling (SDM) [@hijmans2013].
However, by default, the method considers all points for sampling the training and test sets.

@@

The "Spatial CV" method after [@russ2010; @brenning2012] and implemented in package \pkg{sperrorest} uses the coordinates of all observations to create clusters in the spatial domain.
$k$-means clustering has been proposed as a clustering method.
To avoid confusion with other spatial partitioning methods, we refer to this as coordinate-based clustering, or as `spcv_coords` in \pkg{mlr3spatiotempcv}.

This approach is very versatile as it adapts to irregularly-shaped study areas and ensures that exactly $k$ partitions are created.
Nevertheless, despite the random selection of initial cluster centers, repeated partitionings may in some cases be nearly identical. Also, $k$-means clustering may be less suitable for data sets with existing clusters of points and/or with distant isolated sample locations.

```{r coords, fig.cap="Method 'Spatial CV' showing k-means clustering based on coordinates.", out.width="50%", fig.pos="ht"}
resampling_coords = rsmp("spcv_coords", folds = 5)$instantiate(task)
